Make any stupid plot you want

### 2021-01-01 draft/sections/abstract.tex
naturalness

Potentially the most important property of a data visualization is its expressiveness.  We propose to formalize expressiveness via a topological artist model (TAM). In this model, data and graphics can be viewed as sections of fiber bundles; this allows for (1) decomposing the translation of data fields (variables) into visual channels via an equivariant map on the fibers and (2) a topology-preserving map of the base spaces that translates the dataset connectivity into graphical elements. Furthermore, TAM supports an algebraic sum operation such that more complex visualizations can be built from simple ones. We illustrate the application of the model through case studies of a scatter plot, line plot, and heatmap. We show that TAM can be implemented with a small prototype.

>>>>

 A critical aspect of data visualization is that the graphical representation of data match the properties of the data, as stated in Norman's Naturalness Principal. One failure for numerical data is the Lie Factor, another for categorical data is ordering not being preserved. In this work, we propose that the mathematical notion of equivariance formalizes the naturalness principle. We developed a topological artist model (TAM) in which data and graphics can be viewed as sections of fiber bundles. This model allows for (1) decomposing the translation of data fields (variables) into visual channels via an equivariant map on the fibers and (2) a topology-preserving map of the base spaces that translates the dataset connectivity into graphical elements. Furthermore, TAM supports an algebraic sum operation such that more complex visualizations can be built from simple ones. We illustrate the application of the model through case studies of a scatter plot, line plot, and heatmap. We show that TAM can be implemented with a small prototype.
